---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2019"
output: 
  md_document:
    variant: markdown_github
---


# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment. Feel free to ignore the rest (although pause behavior looks interesting, if you check my article).

```{r}
# load libraries
pacman::p_load(pacman,tidyverse,metafor,lmerTest, dplyr)

# read file into a dataframe
crazyData <- readxl::read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")

# subsetting control data
behavedData <- crazyData %>% select("StudyID",TYPE_OF_TASK,DIAGNOSIS,SAMPLE_SIZE_HC,SAMPLE_SIZE_SZ,PITCH_F0_HC_M,PITCH_F0_HC_SD,PITCH_F0_SZ_M, PITCH_F0_SZ_SD)

# subsetting schizophrenia data
behavedDataSD <- crazyData %>% select("StudyID",TYPE_OF_TASK,DIAGNOSIS,SAMPLE_SIZE_HC,SAMPLE_SIZE_SZ,PITCH_F0SD_HC_M,PITCH_F0SD_HC_SD,PITCH_F0SD_SZ_M,PITCH_F0SD_SZ_SD)

# deleting NA's (which downsamples our dataset for metaanalysis, because studies use different measurements)
behavedData <- na.omit(behavedData)
behavedDataSD <- na.omit(behavedDataSD)

# set variables as a factor, introducing levels
behavedData <- behavedData %>% 
mutate_at(c("TYPE_OF_TASK","StudyID"),as.factor)

```


```{r}

# calculating effect size for mean meassures dataframe
behavedData <- metafor::escalc("SMD",
n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC,
m1i = PITCH_F0_SZ_M, m2i = PITCH_F0_HC_M,
sd1i = PITCH_F0_SZ_SD, sd2i = PITCH_F0_HC_SD,data=behavedData)

# set and examine factor levels
behavedData$TYPE_OF_TASK <- behavedData$TYPE_OF_TASK %>% as.factor()
behavedData$TYPE_OF_TASK %>% levels()

# defining a simple model predicting effect size with studies weighted by their squared standard deviation
simpleModel <- lmer(yi ~ 1 + (1|StudyID), behavedData, weights = 1/vi, REML=F,
control = lmerControl(
check.nobs.vs.nlev="ignore",
check.nobs.vs.nRE="ignore"))

# defining a model predicting effect size from type of task with studies weighted by their squared standard deviation
intermediateModel <- lmer(yi ~ 1 + TYPE_OF_TASK + (1|StudyID), behavedData, weights = 1/vi, REML=F,
control = lmerControl(
check.nobs.vs.nlev="ignore",
check.nobs.vs.nRE="ignore"))

# summarize statistical analysis of healthy controls
summary(simpleModel)
summary(intermediateModel)

# defining a model using the rma (regression meta analysis) function that fits the meta-analytic models via linear mixed effects models
metaModel <- rma(yi, vi, data = behavedData, slab=StudyID)


#adding fixed effects
metaFixedModel <- rma(yi, vi, data = behavedData,mods = cbind(TYPE_OF_TASK), slab=StudyID,weighted=T)



# summarize models
summary(metaModel)
summary(metaFixedModel)


# creating forest plots to visualize the standardized mean difference of the studies in the meta analysis

forest(metaModel, main = "Mean Model")
forest(metaFixedModel, main="Mean Model with Fixed Effect = Task")


```
```{r}
# calculating effect size for standard deviation 
behavedDataSD <- metafor::escalc("SMD",
n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC,
m1i = PITCH_F0SD_SZ_M, m2i = PITCH_F0SD_HC_M,
sd1i = PITCH_F0SD_HC_SD, sd2i = PITCH_F0SD_SZ_SD,data=behavedDataSD)

# set and examine factor levels
behavedDataSD$TYPE_OF_TASK <- behavedDataSD$TYPE_OF_TASK %>% as.factor()
behavedDataSD$TYPE_OF_TASK %>% levels()

# defining a simple model predicting effect size with studies weighted by their squared standard deviation
simpleModelSD <- lmer(yi ~ 1 + (1|StudyID), behavedDataSD, weights = 1/vi, REML=F,
control = lmerControl(
check.nobs.vs.nlev="ignore",
check.nobs.vs.nRE="ignore"))

# defining a model predicting effect size from type of task with studies weighted by their squared standard deviation
intermediateModelSD <- lmer(yi ~ 1 + TYPE_OF_TASK + (1|StudyID), behavedDataSD, weights = 1/vi, REML=F,
control = lmerControl(
check.nobs.vs.nlev="ignore",
check.nobs.vs.nRE="ignore"))

# summarize statistical analysis of healthy controls
summary(simpleModelSD)
summary(intermediateModelSD)

# defining a model using the regression meta analysis function 
metaModelSD <- rma(yi, vi, data = behavedDataSD, slab=StudyID)

#adding Fixed effects
metaFixedModelSD <- rma(yi, vi, data = behavedDataSD,mods = cbind(TYPE_OF_TASK), slab=StudyID,weighted=T)


# summarize models
summary(metaModelSD)

summary(metaFixedModelSD)


# creating a forest plot to visualize the standardized mean difference of the studies in the meta analysis

forest(metaModelSD, main = "SD Model")

forest(metaFixedModelSD, main = "SD Model, Fixed effect = Task")



```


2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r}
#### Create separate calculations for our own study assignment 3 ####
# Read data from previous study into data frame
ownData <- read.csv("assignment3.csv")

# Subset dataframe by selecting diagnosis, mean, sd, and subject id.
ownData <- ownData %>% select(Diagnosis, mean, sd, Subject)

# Convert Diagnosis from levels to characters
ownData$Diagnosis <- as.character(ownData$Diagnosis)



# Create a dataframe of means including our study assignment
#SZ
# create schizophrenia data frame
szData <- filter(ownData, Diagnosis == "Schizophrenia")

# calculating number of participants
szNumber <- n_distinct(szData$Subject)
# n = 59

# mean pitch and mean of mean pitch of schizophrenic participants
szMeanVec <- szData %>% group_by(Subject) %>%
  summarize(mean=mean(mean))
szMeanMean <- mean(szMeanVec$mean)

# standard deviation of the mean pitch
szSDMean <- sd(szMeanVec$mean)

# HC
# create a healthy control data frame
hcData <- filter(ownData, Diagnosis == "Control")

 # calculate number of participants
hcNumber <- n_distinct(hcData$Subject)
# n = 59

# find mean pitch and mean of mean pitch of healthy controls
hcMeanVec <- hcData %>% group_by(Subject) %>%
  summarize(mean=mean(mean))
hcMeanMean <- mean(hcMeanVec$mean)

# standard deviation of the mean pitch
hcSDMean <- sd(hcMeanVec$mean)

# create dataframe
ourOwnData <- data.frame(StudyID = "4", TYPE_OF_TASK = "FREE", DIAGNOSIS = "SZ", SAMPLE_SIZE_HC = hcNumber, SAMPLE_SIZE_SZ = szNumber, PITCH_F0_HC_M = hcMeanMean, PITCH_F0_HC_SD = hcSDMean, PITCH_F0_SZ_M = szMeanMean, PITCH_F0_SZ_SD = szSDMean,yi=NA,vi=NA)

# binding data frame into one with previous data
mergedData <- rbind.data.frame(behavedData, ourOwnData)

# creating effectsize and measures of uncertainty
mergedData  <- metafor::escalc("SMD",
n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC,
m1i = PITCH_F0_SZ_M, m2i = PITCH_F0_HC_M,
sd1i = PITCH_F0_SZ_SD, sd2i = PITCH_F0_HC_SD,data=mergedData )


# create a model using effect size and an uncertainty measure
metaModelOwn <- rma(yi, vi, data = mergedData, slab=StudyID)

#adding fixed effects
metaFixedModelOwn <- rma(yi, vi, data = mergedData,cbind("TYPE_OF_TASK"),weighted=T, slab=StudyID)


#looking at models
summary(metaModelOwn)
summary(metaFixedModelOwn)


# create forest plots displaying effectsizes
forest(metaModelOwn, main= "Mean Model")
forest(metaFixedModelOwn, main = "Mean Model, Fixed effect = Task")



# create a data frame of SD's including our study assignment
#SZ
# mean pitch and mean of mean pitch of schizophrenic participants
szSDVec <- szData %>% group_by(Subject) %>%
  summarize(meanSD=mean(sd)) 

# mean and sd of standard deviation vector
szMeanSD <- mean(szSDVec$meanSD)
szSDSD <- sd(szSDVec$meanSD)

# HC
# find mean pitch and mean of mean pitch of healthy controls
hcSDVec <- hcData %>% group_by(Subject) %>%
  summarize(meanSD=mean(sd))

# mean and sd of standard deviation vector
hcMeanSD <- mean(hcSDVec$meanSD)
hcSDSD <- sd(hcSDVec$meanSD)

# create row in dataframe following earlier example
ourOwnDataSD <- data.frame(StudyID = "4", TYPE_OF_TASK = "FREE", DIAGNOSIS = "SZ", SAMPLE_SIZE_HC = hcNumber, SAMPLE_SIZE_SZ = szNumber, PITCH_F0SD_HC_M = hcMeanSD, PITCH_F0SD_HC_SD = hcSDSD, PITCH_F0SD_SZ_M = szMeanSD, PITCH_F0SD_SZ_SD = szSDSD,yi=NA,vi=NA)


# binding data frame into one with previous data
mergedDataSD <- rbind.data.frame(behavedDataSD, ourOwnDataSD)

# creating effectsize and measures of uncertainty
mergedDataSD <- metafor::escalc("SMD",
n1i = SAMPLE_SIZE_SZ, n2i = SAMPLE_SIZE_HC,
m1i = PITCH_F0SD_SZ_M, m2i = PITCH_F0SD_HC_M,
sd1i = PITCH_F0SD_HC_SD, sd2i = PITCH_F0SD_SZ_SD,data=mergedDataSD)


# create a model using effect size and an uncertainty measure
metaModelOwnSD <- rma(yi, vi, data = mergedDataSD, slab=StudyID)

#addign fixed effects
metaFixedModelOwnSD <- rma(yi, vi, data = mergedDataSD,mods = cbind(TYPE_OF_TASK), slab=StudyID,weighted=T)

# summarize models
summary(metaModelOwnSD)

summary(metaFixedModelOwnSD)

# create forest plots displaying effectsizes
forest(metaModelOwnSD, main = "SD Model")

forest(metaFixedModelOwnSD, main = "SD Model, Fixed effect = Task")

```

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r}
# create two funnel plots in order to investigate publication bias
funnel(metaModel, main = "Mean Model")
funnel(metaModelSD, main = "SD Model")

# fixed effects
funnel(metaFixedModel, main = "Mean Model, Fixed Effect = Task")
funnel(metaFixedModelSD, main = "SD Model, Fixed Effect = Task")

# testing for influential studies using influence plots
# meta model with means
influence <- influence(metaModel)
plot(influence)
# meta model with sdandard deviation
influenceSD <- influence(metaModelSD)
plot(influenceSD)



#fixed effects
influenceFixed <- influence(metaFixedModel)
plot(influenceFixed)
# meta model with sdandard deviation
influenceSDFixed <- influence(metaFixedModelSD)
plot(influenceSDFixed)


# Extra tests
# funnel
regtest(metaModel) # test for funnel plot asymmetry: z = 1.5798, p = 0.1142
regtest(metaModelSD) # test for funnel plot asymmetry: z = 0.2004, p = 0.8412

# kendall's tau
ranktest(metaModel) # k tau = 0.20, p = 0.72
ranktest(metaModelSD) # k tau = -0.28, p = 0.69
# tau 2
metaModel$tau2 # 0.071
metaModelSD$tau2 # 1.30

# calculate I square from the two meta models
metaModel$I2
# 50.29
metaModelSD$I2
# 95.37


# Extra tests
# funnel
regtest(metaFixedModel) # test for funnel plot asymmetry: z = 1.1308, p = 0.2581
regtest(metaFixedModelSD) # test for funnel plot asymmetry: z = 0.3027, p = 0.7621

# kendall's tau
ranktest(metaFixedModel) #Kendall's tau = 0.2000, p = 0.7194
ranktest(metaFixedModelSD) #Kendall's tau = -0.2762, p = 0.1686
# tau^2
metaFixedModel$tau2 # 0.09239298
metaFixedModelSD$tau2 #1.35172

# calculate I square from the two meta models
metaFixedModel$I2
# 55.33512
metaFixedModelSD$I2
# 95.56817



```


## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: PITCH_F0M and PITCH_F0SD group of variables are what you need
    * Hint: Make sure you read the comments in the columns: `pitch_f0_variability`, `frequency`, `Title`,  `ACOUST_ANA_DESCR`, `DESCRIPTION`, and frma`COMMENTS`
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
